\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics

### 2017-04-25 journalSwarmControl/Introduction.tex
Some of Dylan's comments are affected

This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a
single journal paper.
  \cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.  This paper presents three years of results from new games at SwarmControl.net.  For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parametric studies.  All hardware validation experiments are new.

@@

\begin{figure}
%\centering
\renewcommand{\figwid}{0.3\columnwidth}
\begin{overpic}[width =\figwid]{VaryNum.pdf}	\put(10,80){\textbf{a} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}	\put(10,80){\textbf{b} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}\\
\\
\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
\caption{\label{fig:5experiments}
Screenshots from our online experiments controlling multi-particle systems with limited, global control.
\textbf{(a)} Varying the number of particles from 1-500
\textbf{(b)} Comparing 4 levels of visual feedback 
\textbf{(c)} Varying noise from 0 to 200\% of control authority
\textbf{(d)} Controlling the position of 1 to 10 particles
\textbf{(e)} Comparing 3 control architectures to assemble
\textbf{(f)} Comparing 3 control architectures to forage.
}
\end{figure}

>>>>

%%%% Move it to the cover letter.
%This paper combines the content of two preliminary conference papers, extending their substance and providing full details in a single journal paper.
 % \cite{swarmcontrol2013} covered the first three months of SwarmControl.net experiments, and \cite{ShahrokhiIROS2015} presented simulations of object manipulation.  This paper presents three years of results from new games at SwarmControl.net.  For object manipulation, this paper presents robust new algorithms for manipulation, path planning, and obstacle avoidance, and a rich set of parametric studies.  All hardware validation experiments are new.

@@

\begin{figure}
%\centering
\renewcommand{\figwid}{0.3\columnwidth}
\begin{overpic}[width =\figwid]{VaryNum.pdf}	\put(10,80){\textbf{a} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}	\put(10,80){\textbf{b} }\end{overpic}~
\begin{overpic}[width =\figwid]{VaryNoise.pdf}	\put(10,80){\textbf{c} }\end{overpic}\\
%\\
%\begin{overpic}[width =\figwid]{ControlPos.pdf}	\put(10,80){\textbf{d} }\end{overpic}~
%\begin{overpic}[width =\figwid]{VaryControl.pdf}	\put(10,80){\textbf{e} }\end{overpic}~
%\begin{overpic}[width =\figwid]{VaryForage.pdf}	\put(10,80){\textbf{f} }\end{overpic}
\caption{\label{fig:5experiments}
Screenshots from our online experiments controlling multi-particle systems with limited, global control.
\textbf{(a)} Varying the number of particles from 1-500
\textbf{(b)} Comparing 4 levels of visual feedback 
\textbf{(c)} Varying noise from 0 to 200\% of control authority
%\textbf{(d)} Controlling the position of 1 to 10 particles
%\textbf{(e)} Comparing 3 control architectures to assemble
%\textbf{(f)} Comparing 3 control architectures to forage.
}
\end{figure}

### 2017-04-25 journalSwarmControl/expResults.tex
Some of Dylan's comments are affected

We designed six experiments to investigate human control of large swarms for manipulation tasks.  Screenshots of each experiment are shown in Fig.~\ref{fig:5experiments}.  Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks. The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.  We recorded the completion time and the participant ID for each successful trial.  
%As Fig.~\ref{fig:Learning} shows, one-third of all participants played only a single game.  Still, many played multiple games, and their decreasing completion times demonstrates their skills improved.

\paragraph{Varying number}
\begin{figure}
\begin{overpic}[width = 0.9\columnwidth]{ResVaryNum.pdf}\end{overpic}
%\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
\caption{
\label{fig:ResVaryNu}Data from \emph{Varying Number} using particles to push an object through a maze to a goal location. 
% Best-fit linear and quadratic lines are overlaid for comparison. 
%(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
}
\end{figure}

@@

\paragraph{Varying visualization}
\begin{figure}[b!]
\renewcommand{\figwid}{0.24\columnwidth}
\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
\vspace{-.5em}
\caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}. This experiment challenges players to quickly steer 100 particles (blue discs) to push an object (green hexagon) into a goal region. We record the completion time and other statistics.
%\vspace{-1em}
}
\end{figure}

@@

\paragraph{Varying noise}
%Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
This experiment varied the strength of %these 
disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).

We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the particles move erratically.  Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise.

@@

\paragraph{Position control}
This online experiment examined how completion time scales with the number of particles $n$. Using a single square obstacle, users arranged $n\in[1,10]$ particles into a specified goal pattern.  The goal pattern formed a block  {\sffamily A} with 10 particles, and lesser numbers of particles used a subset of these goal positions. Our hypothesis was that completion time would increase linearly with the number of particles, as with our position control algorithm in \cite{Becker2013b}.  Our results roughly corroborate this, as shown in Fig.~\ref{fig:ResVaryNoisePosition}.b.  Though the number of particles presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.

@@

\paragraph{Varying control--Assembly}
Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.

We compared attractive and repulsive control with the global control used for the other experiments. The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.\ \cite{Khalil2013} and biological models, e.g. \cite{goodrich2012types}. STMs can be used to arrange atoms and make small assemblies, as described in \cite{avouris1995manipulation}. An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules. In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
The experiment challenged players to assemble a three-block pyramid with a swarm of 16 particles.

%\todo{describe controller model here, use an equation}

The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}a: attractive control was the fastest, followed by global control, with repulsive control a distant last.  The median time using repulsive control was four times longer than with attractive control.
Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.

%%\begin{figure}
%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
%%\caption{
%%\label{fig:attractiveForce}
%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
%%}
%%\end{figure}

\begin{figure}[b!]
\renewcommand{\figwid}{0.45\columnwidth}
\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
\begin{overpic}[width =\figwid]{ResVaryForage.pdf}\end{overpic}
\vspace{-.5em}
\caption{\label{fig:ResVaryControl} Completion time depends on both the task and the control type. Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid. Right: in the foraging test, global control resulted in the shortest completion time and attractive the longest.
%\vspace{-1em}
}
\end{figure}

\paragraph{Varying control--Foraging}
Collecting and delivering resources is necessary for drug delivery.
This experiment also compared attractive and repulsive control with the global control used for the other experiments. The experiment challenged players to collect particles using a swarm of 100 particles and return the particles to a home region. The particles encapsulate the particles on contact.

%\todo{describe controller model here, use an equation}

The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.b: global control was the fastest, followed by repulsive control, with attractive control last.  Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $2.96\times10^{-6}$.

>>>>

We designed six experiments to investigate human control of large swarms for manipulation tasks.  Screenshots of representative experiment are shown in Fig.~\ref{fig:5experiments}.  Each experiment examined the effects of varying a single parameter: population of particles for manipulation, four levels of visual feedback, different levels of Brownian noise, position control with 1 to 10 particles, and three control architectures for both assembly and foraging tasks. The users could choose which experiment to try, and our architecture randomly assigned a parameter value for each trial.  We recorded the completion time and the participant ID for each successful trial.  
%As Fig.~\ref{fig:Learning} shows, one-third of all participants played only a single game.  Still, many played multiple games, and their decreasing completion times demonstrates their skills improved.

\subsection{Varying number}
\begin{figure}
\begin{overpic}[width = 0.9\columnwidth]{ResVaryNum.pdf}\end{overpic}
%\begin{overpic}[width = 0.48\columnwidth]{measureLearning.pdf}\end{overpic}
\caption{
\label{fig:ResVaryNu}Data from \emph{Varying Number} using particles to push an object through a maze to a goal location. 
% Best-fit linear and quadratic lines are overlaid for comparison. 
%(Right) Skill improves as players retry tests using data from \emph{Varying Number}.
}
\end{figure}

@@

\subsection{Varying visualization}
\begin{figure}[b!]
\renewcommand{\figwid}{0.24\columnwidth}
\begin{overpic}[width =\figwid]{VaryVisFS.pdf}\put(20,15){Full-state}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisCH.pdf}\put(10,15){Convex-hull}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMV.pdf}\put(10,15){Mean + var}\end{overpic}
\begin{overpic}[width =\figwid]{VaryVisMe.pdf}\put(30,15){Mean}\end{overpic}
\vspace{-.5em}
\caption{\label{fig:Visualization}Screenshots from task \emph{Vary Visualization}. This experiment challenges players to quickly steer 100 particles (blue discs) to push an object (green hexagon) into a goal region. We record the completion time and other statistics.
%\vspace{-1em}
}
\end{figure}

@@

\subsection{Varying noise}
%Micro-particles are affected by random collisions with molecules. The effect of these collisions is called Brownian motion.
This experiment varied the strength of %these 
disturbances to study how noise affects human control of large swarms. Noise was applied at every timestep:
\begin{align*}
\dot{x}_i &= u_x + m_i\cos(\psi_i)\\
 \dot{y}_i &= u_y + m_i\sin(\psi_i).
 \end{align*}
$m_i,\psi_i$ were uniformly IID, with $m_i\in[0,M]$ and $\psi_i\in[0,2\pi]$. $M$ was a constant for each trial ranging from 0 to 200\% of the maximum control power ($u_{max}$).

We hypothesized 200\% noise was the largest a human could be expected to control---at 200\% noise, the particles move erratically.  Disproving our hypothesis, the results in Fig.~\ref{fig:ResVaryNoisePosition}.a show only a 40\% increase in completion time for the maximum noise. This indicates swarm control is robust to IID noise.

@@

%\paragraph{Position control}
%This online experiment examined how completion time scales with the number of particles $n$. Using a single square obstacle, users arranged $n\in[1,10]$ particles into a specified goal pattern.  The goal pattern formed a block  {\sffamily A} with 10 particles, and lesser numbers of particles used a subset of these goal positions. Our hypothesis was that completion time would increase linearly with the number of particles, as with our position control algorithm in \cite{Becker2013b}.  Our results roughly corroborate this, as shown in Fig.~\ref{fig:ResVaryNoisePosition}.b.  Though the number of particles presented to game players is uniformly distributed, larger $n$ are more difficult, and the number of successful experiments drops steadily as $n$ increases.
%

@@

%\paragraph{Varying control--Assembly}
%Ultimately, we want to use swarms of particles to build things. This experiment compared different control architectures modeled after real-world devices.
%
%We compared attractive and repulsive control with the global control used for the other experiments. The attractive and repulsive controllers were loosely modeled after scanning tunneling microscopes (STM), but also apply to magnetic manipulation, e.g.\ \cite{Khalil2013} and biological models, e.g. \cite{goodrich2012types}. STMs can be used to arrange atoms and make small assemblies, as described in \cite{avouris1995manipulation}. An STM tip is charged with electrical potential, and used to repel like-charged or to attract differently-charged molecules. In contrast, the global controller uses a uniform field (perhaps formed by parallel lines of differently-charged conductors) to pull molecules in the same direction.
%The experiment challenged players to assemble a three-block pyramid with a swarm of 16 particles.
%
%\todo{describe controller model here, use an equation}

%The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}a: attractive control was the fastest, followed by global control, with repulsive control a distant last.  The median time using repulsive control was four times longer than with attractive control.
%Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $3.37\times10^{-32}$.
%
%%\begin{figure}
%%\begin{overpic}[width = 0.39\columnwidth]{attractiveForce.pdf}\end{overpic}
%%\begin{overpic}[width = 0.6\columnwidth]{understandingForces.pdf}\end{overpic}
%%\caption{
%%\label{fig:attractiveForce}
%%Attraction and repulsion with a point source distance $h$ above the plane and $r$ from the robot. 
%%}
%%\end{figure}

%\begin{figure}[b!]
%\renewcommand{\figwid}{0.45\columnwidth}
%\begin{overpic}[width =\figwid]{ResVaryControl.pdf}\end{overpic}
%\begin{overpic}[width =\figwid]{ResVaryForage.pdf}\end{overpic}
%\vspace{-.5em}
%\caption{\label{fig:ResVaryControl} Completion time depends on both the task and the control type. Left: Attractive control resulted in the shortest completion time and repulsive the longest for building a three-block pyramid. Right: in the foraging test, global control resulted in the shortest completion time and attractive the longest.
%%\vspace{-1em}
%}
%\end{figure}
%
%\paragraph{Varying control--Foraging}
%Collecting and delivering resources is necessary for drug delivery.
%This experiment also compared attractive and repulsive control with the global control used for the other experiments. The experiment challenged players to collect particles using a swarm of 100 particles and return the particles to a home region. The particles encapsulate the particles on contact. 
%
%%\todo{describe controller model here, use an equation}
%
%The results were conclusive, as shown in Fig.~\ref{fig:ResVaryControl}.b: global control was the fastest, followed by repulsive control, with attractive control last.  Using ANOVA analysis, we reject the null hypothesis that all controllers are equivalent, with $p$-value $2.96\times10^{-6}$.
%
%

### 2017-04-25 journalSwarmControl/experiment.tex
Some of Dylan's comments are affected

To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
  Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}: a function that assigns a control input to every possible state.
 At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.

 A corresponding reward function gives a high reward to the goal state,  
 a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to a state with an obstacle} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 $\mathbf{u}$ is the action and $x$ is the current state.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j| \mathbf{u},x)].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| \mathbf{u},x_j)] \label{eq:ValueIteration} \nonumber\\
\text{end}& 
\end{align}
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}b.
\begin{align}
F_{att} &= -\zeta \Delta \rho / \rho \\
%\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right) \\
%\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
 F_{rep} &=  \left\{
\begin{array}{ll}
      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\&~\phi <  \theta \\
      0, & \textrm{ otherwise} \\
\end{array} 
\right.\\
F_{pot} &= F_{att} + F_{rep} \label{eq:potentialfield}
\end{align}

@@

The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, often because workspace obstacles made some particles unable to reach the object. This failure rate increases if  object weight increases or ground-robot friction increases. The mean and covariance calculations \eqref{eq:meanVar} included all robots in the workspace. Robots that cannot reach the object due to obstacles skew these calculations. The state machine in Fig.\ \ref{fig:Region}a solves this problem by creating two states for the maze: either main or transfer. Each state has a set of regions representing a discretized visibility polygon. Whenever the object crosses a region boundary the state toggles. The main regions are generated by extending obstacles until they meet another obstacle. The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.

>>>>

To solve both these problems, this paper models object movement as a Markov Decision Process (MDP) with non-deterministic movement.  
  Value iteration,  as described in \cite{Thrun2005}, is used to learn an \emph{optimal policy}.
   At each state the object can be commanded to move in one of eight directions with a small probability of moving in a wrong direction.

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to a state with an obstacle} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $\mathbf{u}$ is the action and $x$ is the current state.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}   [ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j|x, \mathbf{u})].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_j) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_j) = \gamma \max_{\mathbf{u}} [r(x_j,\mathbf{u}) + \sum\limits_{i=1}^N \hat{V}(x_i) p(x_i| x_j,\mathbf{u})] \label{eq:ValueIteration} \nonumber\\
\text{end}& 
\end{align}
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation of this algorithm is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

Unfortunately, when the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach inspired by \cite{spong2008robot} that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object.
The repulsive potential field is centered at the object's COM and is active for a radius $\rho_0$, but is implemented only when the swarm mean is within $\theta$ of the desired direction of motion $\mathbf{D}(\mathbf{b})$ as shown in Fig.~\ref{fig:potentialField}b.
\begin{align}
F_{att} &= -\zeta \Delta \rho / \rho, \\
%\mathbf{u} &= [\bar{x},\bar{y}] - \mathbf{b} \nonumber \\
\phi &=\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot ( [\bar{x},\bar{y}] - \mathbf{b}) }{\norm{\mathbf{D}(\mathbf{b})} \norm{ [\bar{x},\bar{y}] - \mathbf{b}} } \right), \\
%\theta &= \mathrm{angularDistance}\left( \alpha, \beta \right)\nonumber \\
 F_{rep} &=  \left\{
\begin{array}{ll}
      \eta( 1/\rho- 1/\rho_0) \frac{1}{\rho^2} \Delta \rho, ~& \rho\leq \rho_0~\wedge~\phi <  \theta \\
      0, & \textrm{ otherwise} \\
\end{array},
\right.\\
F_{pot} &= F_{att} + F_{rep} \label{eq:potentialfield}.
\end{align}

@@

The variance controller in Alg.~\ref{alg:MeanVarianceControl} is a greedy algorithm that is susceptible to outliers. 
The controller in \cite{ShahrokhiIROS2015} failed in $14\%$ trials, some particles were unable to reach the object because workspace obstacles were blocking them. This failure rate increases if  object weight increases or ground-robot friction increases. The mean and covariance calculations \eqref{eq:meanVar} included all robots in the workspace. Robots that cannot reach the object due to obstacles skew these calculations. The state machine in Fig.\ \ref{fig:Region}a solves this problem by creating two states for the maze: either main or transfer. Each state has a set of regions representing a discretized visibility polygon. Whenever the object crosses a region boundary the state toggles. The main regions are generated by extending obstacles until they meet another obstacle. The transfer regions are perpendicular to obstacle boundaries, and act as a buffer between two main regions.

### 2017-04-25 journalSwarmControl/relatedWork.tex
Some of Dylan's comments are affected

 \paragraph{Control---global inputs:}
 Many micro- and nanorobotic systems, see \cite{Tottori2012,Shirai2005,Chiang2011,Donald2006,Donald2008,Takahashi2006,Floyd2011,Diller2013,Frutiger2008,Peyer2013}
   rely on global inputs, where each robot receives an exact copy of the control signal.  Our experiments follow this global model.
   %Two reasonable questions are ``What tasks are possible with many robots, all under uniform control inputs?'' and ``What tasks are impossible with many robots, all under uniform control inputs?''

 \paragraph{Sensing---large populations:}
$n$ differential-drive robots in a 2D workspace require $3n$ state variables. Even holonomic robots require $2n$ state variables. Numerous methods exist for measuring this state in microrobotics.  These solutions use computer vision systems to sense position and heading angle, with corresponding challenges of handling missed detections and image registration between detections and robots.  These challenges are increased at the nanoscale where sensing competes with control for communication bandwidth.   We examine control when the operator has access to partial feedback, including only the first and/or second moments of a population's position, or only the convex-hull containing the robots.

\paragraph{Computation---calculating the control law:}
In our previous work the controllers required at best a summation over all the robot states, see \cite{Becker2012k} and at worst a matrix inversion, see \cite{Becker2012}. 
These operations become intractable for large populations of robots. By focusing on \emph{human} control of large robot populations, we accentuate computational difficulties because the controllers are implemented by the unaided human operator. % We present an approach that, for many tasks, bypasses these problems due to large populations by allowing the user to command the entire population as a single unit.  For position control we cannot bypass this problem, but we provide an algorithm that scales linearly in the number of robots.

@@

  Most humans are able to, with practice, steer a swarm of robots controlled by a global input. Prior to our paper, no algorithm existed. Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}.This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.

 \cite{ayanian2014controlling} designed a user interface enabling an operator to maneuver a swarm of robots through a cluttered workspace by specifying the bounding prism for the swarm and then translating or scaling this prism. Our paper shares the concept of a global control input, but our robots have no onboard computation and cannot track a virtual boundary.

\cite{Jr2004} studied human \emph{fanout}, the number of robots a single human user could directly control.  %Studied as a key aspect of human-robot interaction, 
They postulated that the optimal number of robots was approximately  the autonomous time  divided by the interaction time required by each robot.  Their sample problem involved a multi-robot search task, where users could assign goals to robots.  Their user interaction studies with simulated planar robots  indicated a \emph{fanout plateau} of about 8 robots, with diminishing returns for more robots.   They hypothesized the location of this plateau is highly dependent on the underlying task.  Indeed, our paper indicates there are tasks without plateaus. % (Lemmings is an example of this as well, for systems with autonomy
Their research investigated robots with 3 levels of autonomy.  We use robots without autonomy, corresponding with their first-level robots.

@@

Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.  Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.  Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in\S \ref{sec:VarianceControl}.

>>>>

 \paragraph{Control---global inputs}
 Many micro- and nanorobotic systems, see \cite{Tottori2012,Shirai2005,Chiang2011,Donald2006,Donald2008,Takahashi2006,Floyd2011,Diller2013,Frutiger2008,Peyer2013}
   rely on global inputs, where each robot receives an exact copy of the control signal.  Our experiments follow this global model.
   %Two reasonable questions are ``What tasks are possible with many robots, all under uniform control inputs?'' and ``What tasks are impossible with many robots, all under uniform control inputs?''

 \paragraph{Sensing---large populations}
$n$ differential-drive robots in a 2D workspace require $3n$ state variables. Even holonomic robots require $2n$ state variables. Numerous methods exist for measuring this state in microrobotics.  These solutions use computer vision systems to sense position and heading angle, with corresponding challenges of handling missed detections and image registration between detections and robots.  These challenges are increased at the nanoscale where sensing competes with control for communication bandwidth.   We examine control when the operator has access to partial feedback, including only the first and/or second moments of a population's position, or only the convex-hull containing the robots.

\paragraph{Computation---calculating the control law}
In our previous work the controllers required at best a summation over all the robot states, see \cite{Becker2012k} and at worst a matrix inversion, see \cite{Becker2012}. 
These operations become intractable for large populations of robots. By focusing on \emph{human} control of large robot populations, we accentuate computational difficulties because the controllers are implemented by the unaided human operator. % We present an approach that, for many tasks, bypasses these problems due to large populations by allowing the user to command the entire population as a single unit.  For position control we cannot bypass this problem, but we provide an algorithm that scales linearly in the number of robots.

@@

  Most humans are able to, with practice, steer a swarm of robots controlled by a global input. Prior to our paper, no algorithm existed. Using human input to learn how to control a dynamic system is a line of research with a rich history \cite{argall2009, abbeel2010}. This paper exploits insights gained from ``swarmcontrol.net", particularly the fact that having a swarm's mean and variance is sufficient for object manipulation through an obstacle field.

 A user interface enabling an operator to maneuver a swarm of robots through a cluttered workspace by specifying the bounding prism for the swarm and then translating or scaling this prism is designed in \cite{ayanian2014controlling}. Our paper shares the concept of a global control input, but our robots have no onboard computation and cannot track a virtual boundary.

Human \emph{fanout}, the number of robots a single human user could directly control is studied in \cite{Jr2004}.  %Studied as a key aspect of human-robot interaction, 
They postulated that the optimal number of robots was approximately  the autonomous time  divided by the interaction time required by each robot.  Their sample problem involved a multi-robot search task, where users could assign goals to robots.  Their user interaction studies with simulated planar robots  indicated a \emph{fanout plateau} of about 8 robots, with diminishing returns for more robots.   They hypothesized the location of this plateau is highly dependent on the underlying task.  Indeed, our paper indicates there are tasks without plateaus. % (Lemmings is an example of this as well, for systems with autonomy
Their research investigated robots with 3 levels of autonomy.  We use robots without autonomy, corresponding with their first-level robots.

@@

Our $n$-robot system with 2 control inputs and 4$n$ states is inherently under-actuated, and superficially bears resemblance to compliant, under-actuated manipulators.  Our swarms conform to the object to be manipulated, but lack the restoring force provided by flexures in \cite{odhner2014compliant} or silicone in \cite{deimel2014novel}.  Our swarms tend to disperse and so to regroup them we require artificial forces like the variance control primitives in \S \ref{sec:VarianceControl}.
